---
title: "Incidence estimates for Takayasu arteritis in England by Demographic Group: repeat of previous modelling to harvest mutually adjusted age and sex coefficients"
author: "SJ Lax"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
  pdf_document: default
---

### Setup

```{r setup}
#install packages at first use
#install.packages("tidyverse", repos = "https://packages.sde.digital.nhs.uk/repository/cran-mirror/", dependencies = TRUE)
#install.packages("knitr", repos = "https://packages.sde.digital.nhs.uk/repository/cran-mirror/", dependencies = TRUE)
#install.packages("kableExtra", repos = "https://packages.sde.digital.nhs.uk/repository/cran-mirror/", dependencies = TRUE)
#install.packages("MASS", repos = "https://packages.sde.digital.nhs.uk/repository/cran-mirror/", dependencies = TRUE)
#install.packages("lmtest", repos = "https://packages.sde.digital.nhs.uk/repository/cran-mirror/", dependencies = TRUE)

#setup markdown
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(tidyverse)
library(knitr)
library(kableExtra)
library(MASS)
library(lmtest)

setwd("D:/PhotonUser/My Files/Home Folder/ccu075_01")

# Get cohort data
incidence_cohort_England <- read.csv2("ccu075_01_M314_cohort_20240925.csv")

merged <- read.csv("ccu075_01_denominators_i_age_group_sex_ethnicity_20241010.csv")

# Set parameters for analyses
CI_level <- 0.95
rate_multiplier <- 1000000

#Compile cases and denominator dataset where age is grouped
grouped_demographics <- incidence_cohort_England %>%
  group_by(Eth_grouped, sex, age_group) %>%
  summarise(cases = n()) %>%
  ungroup() %>%
  rename(ethnicity_broad = Eth_grouped)

merged_grouped <- full_join(grouped_demographics, merged, by=c("ethnicity_broad", "sex", "age_group")) 
```

### Checking crude incidence rate ratios: sex

```{r incidence_odds_adj_sex, message=FALSE}
# Resolves 0 counts to 1 to enable code to parse
adjusted_inc_rates_by_ethnicity <- merged_grouped %>%
  mutate(Cases_for_calculations = ifelse(cases == 0 | is.na(cases), 1, cases)) 

#redo variable type for ungrouped dataset
adjusted_inc_rates_by_ethnicity$sex <- factor(adjusted_inc_rates_by_ethnicity$sex, levels=c("Male", "Female"))                     

poisson.model_sex <- glm(formula = Cases_for_calculations ~ sex, 
                         offset = log(combined_denominator),
                         family = poisson(link="log"), 
                         data = adjusted_inc_rates_by_ethnicity)

exp(coef(poisson.model_sex)) # consistent with previous estimates; not expected to be identical because of different process for generating Cases_for_calculations
```

### Checking crude incidence rate ratios: age

```{r incidence_odds_adj_age, message=FALSE}
adjusted_inc_rates_by_ethnicity$age_group <- factor(adjusted_inc_rates_by_ethnicity$age_group, levels=c("16-39",
                                                                                        "0-15",
                                                                                        "40-54",
                                                                                        "55-69",
                                                                                        "70-84",
                                                                                        "85-90"))

poisson.model_age <- glm(formula = Cases_for_calculations ~ age_group, 
                         offset = log(combined_denominator),
                         family = poisson(link="log"), 
                         data = adjusted_inc_rates_by_ethnicity)

exp(coef(poisson.model_age)) # consistent with previous estimates; not expected to be equal, as above
```

### Adjusted incidence rate ratios: age and sex

```{r incidence_odds_adj_age, message=FALSE}
poisson.model_sex_by_age <- glm(formula = Cases_for_calculations ~ sex + age_group, 
                         offset = log(combined_denominator),
                         family = poisson(link="log"), 
                         data = adjusted_inc_rates_by_ethnicity)

summary(poisson.model_sex_by_age)

ecm <- exp(coef(poisson.model_sex_by_age))
ci <- exp(confint(poisson.model_sex_by_age))

collated_ORs <- data.frame(rbind(c(Group = "Female (adjusted by age group)", 
                                   OR = round(ecm[str_detect(names(ecm), pattern="Female")][[1]],2), 
                                   Lower_CI = round(ci[str_detect(row.names(ci), pattern="Female"),][[1]],2), 
                                   Upper_CI = round(ci[str_detect(row.names(ci), pattern="Female"),][[2]],2)),
                                 c(Group = "0-15 (adjusted by sex)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="0-15")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="0-15"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="0-15"),][[2]],2)),
                                 c(Group = "40-54 (adjusted by sex)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="40-54")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="40-54"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="40-54"),][[2]],2)),
                                 c(Group = "55-69 (adjusted by sex)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="55-69")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="55-69"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="55-69"),][[2]],2)),
                                 c(Group = "70-84 (adjusted by sex)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="70-84")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="70-84"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="70-84"),][[2]],2)),
                                 c(Group = "85-90 (adjusted by sex)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="85-90")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="85-90"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="85-90"),][[2]],2))
                                 ))
```

### Checking crude incidence rate ratios: ethnicity

```{r incidence_odds_eth_check, message=FALSE}
adjusted_inc_rates_by_ethnicity <- adjusted_inc_rates_by_ethnicity %>%
  filter(ethnicity_broad != "Unknown") 

adjusted_inc_rates_by_ethnicity$ethnicity_broad <- factor(adjusted_inc_rates_by_ethnicity$ethnicity_broad, levels=c("White", 
                                                                                                    "Asian", 
                                                                                                    "Black", 
                                                                                                    "Mixed", 
                                                                                                    "Other"))

poisson.model_ethnicity <- glm(formula = Cases_for_calculations ~ ethnicity_broad, 
                         offset = log(combined_denominator),
                         family = poisson(link="log"), 
                         data = adjusted_inc_rates_by_ethnicity)

exp(coef(poisson.model_ethnicity)) # consistent with previous estimates; not expected to be equal, as above
```

### Adjusted incidence rate ratios: ethnicity, age and sex

```{r incidence_odds_eth, message=FALSE}
poisson.model_ethnicity2 <- glm(formula = Cases_for_calculations ~ sex + ethnicity_broad + age_group, 
                         offset = log(combined_denominator),
                         family = poisson(link="log"), 
                         data = adjusted_inc_rates_by_ethnicity)

summary(poisson.model_ethnicity2)

ecm <- exp(coef(poisson.model_ethnicity2))
ci <- exp(confint(poisson.model_ethnicity2)) # consistent with previous estimates

collated_ORs <- data.frame(rbind(collated_ORs,
                                       c(Group = "Female (adjusted by age group and ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="Female")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="Female"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="Female"),][[2]],2)),
                                 c(Group = "0-15 (adjusted by sex and ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="0-15")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="0-15"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="0-15"),][[2]],2)),
                                 c(Group = "40-54 (adjusted by sex and ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="40-54")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="40-54"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="40-54"),][[2]],2)),
                                 c(Group = "55-69 (adjusted by sex and ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="55-69")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="55-69"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="55-69"),][[2]],2)),
                                 c(Group = "70-84 (adjusted by sex and ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="70-84")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="70-84"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="70-84"),][[2]],2)),
                                 c(Group = "85-90 (adjusted by sex and ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="85-90")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="85-90"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="85-90"),][[2]],2)),
                                 c(Group = "Asian (adjusted by age and sex)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="Asian")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="Asian"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="Asian"),][[2]],2)),
                                       c(Group = "Black (adjusted by age and sex)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="Black")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="Black"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="Black"),][[2]],2)),
                                       c(Group = "Mixed (adjusted by age and sex)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="Mixed")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="Mixed"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="Mixed"),][[2]],2)),
                                       c(Group = "White (adjusted by age and sex)", 
                                         OR = 1, 
                                         Lower_CI = NA, 
                                         Upper_CI = NA),
                                       c(Group = "Other (adjusted by age and sex)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="Other")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="Other"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="Other"),][[2]],2))
                                 )
                           )
```

### Adjusted incidence rate ratios: consideration of interaction between age and ethnicity

```{r incidence_odds_eth_int, message=FALSE}
adjusted_inc_rates_by_ethnicity <- adjusted_inc_rates_by_ethnicity %>%
  mutate(ethnicity_binary = ifelse(ethnicity_broad == "White", 1, 0)) 

## model fit with interaction
poisson.model_ethnicity3 <- glm(formula = Cases_for_calculations ~ sex + ethnicity_binary * age_group, 
                         offset = log(combined_denominator),
                         family = poisson(link="log"), 
                         data = adjusted_inc_rates_by_ethnicity)

summary(poisson.model_ethnicity3)

# Get coefficients
ecm <- exp(coef(poisson.model_ethnicity3))
ci <- exp(confint(poisson.model_ethnicity3))

# Remove interaction coefficients to get only the adjusted age- and sex-specific coefficients for the non-White individuals
other_ecm <- ecm[!grepl(":", names(ecm))]
other_ci <- ci[!grepl(":", rownames(ci)),]

collated_ORs <- data.frame(rbind(collated_ORs,
                                       c(Group = "Female (adjusted by age * ethnicity)", 
                                         OR = round(other_ecm[str_detect(names(other_ecm), pattern="Female")][[1]],2), 
                                         Lower_CI = round(other_ci[str_detect(row.names(other_ci), pattern="Female"),][[1]],2), 
                                         Upper_CI = round(other_ci[str_detect(row.names(other_ci), pattern="Female"),][[2]],2)),
                                 c(Group = "0-15 (non-White only; adjusted by sex)", 
                                         OR = round(other_ecm[str_detect(names(other_ecm), pattern="0-15")][[1]],2), 
                                         Lower_CI = round(other_ci[str_detect(row.names(other_ci), pattern="0-15"),][[1]],2), 
                                         Upper_CI = round(other_ci[str_detect(row.names(other_ci), pattern="0-15"),][[2]],2)),
                                 c(Group = "40-54 (non-White only; adjusted by sex)", 
                                         OR = round(other_ecm[str_detect(names(other_ecm), pattern="40-54")][[1]],2), 
                                         Lower_CI = round(other_ci[str_detect(row.names(other_ci), pattern="40-54"),][[1]],2), 
                                         Upper_CI = round(other_ci[str_detect(row.names(other_ci), pattern="40-54"),][[2]],2)),
                                 c(Group = "55-69 (non-White only; adjusted by sex)", 
                                         OR = round(other_ecm[str_detect(names(other_ecm), pattern="55-69")][[1]],2), 
                                         Lower_CI = round(other_ci[str_detect(row.names(other_ci), pattern="55-69"),][[1]],2), 
                                         Upper_CI = round(other_ci[str_detect(row.names(other_ci), pattern="55-69"),][[2]],2)),
                                 c(Group = "70-84 (non-White only; adjusted by sex)", 
                                         OR = round(other_ecm[str_detect(names(other_ecm), pattern="70-84")][[1]],2), 
                                         Lower_CI = round(other_ci[str_detect(row.names(other_ci), pattern="70-84"),][[1]],2), 
                                         Upper_CI = round(other_ci[str_detect(row.names(other_ci), pattern="70-84"),][[2]],2)),
                                 c(Group = "85-90 (non-White only; adjusted by sex)", 
                                         OR = round(other_ecm[str_detect(names(other_ecm), pattern="85-90")][[1]],2), 
                                         Lower_CI = round(other_ci[str_detect(row.names(other_ci), pattern="85-90"),][[1]],2), 
                                         Upper_CI = round(other_ci[str_detect(row.names(other_ci), pattern="85-90"),][[2]],2))
                                 )
                           )

# Then, to get the age- and sex-specific coefficients for the White individuals, we need use a recoded version of the binary ethnicity variable
adjusted_inc_rates_by_ethnicity <- adjusted_inc_rates_by_ethnicity %>%
  mutate(ethnicity_binary2 = ifelse(ethnicity_broad == "White", 0, 1)) 

## model fit with interaction
poisson.model_ethnicity4 <- glm(formula = Cases_for_calculations ~ sex + ethnicity_binary2 * age_group, 
                         offset = log(combined_denominator),
                         family = poisson(link="log"), 
                         data = adjusted_inc_rates_by_ethnicity)

summary(poisson.model_ethnicity4)

# Get coefficients
ecm <- exp(coef(poisson.model_ethnicity4))
ci <- exp(confint(poisson.model_ethnicity4))

# Remove interaction coefficients to get only the adjusted age- and sex-specific coefficients for the White individuals
white_ecm <- ecm[!grepl(":", names(ecm))]
white_ci <- ci[!grepl(":", rownames(ci)),]

collated_ORs <- data.frame(rbind(collated_ORs,
                                       c(Group = "0-15 (White only; adjusted by sex)", 
                                         OR = round(white_ecm[str_detect(names(white_ecm), pattern="0-15")][[1]],2), 
                                         Lower_CI = round(white_ci[str_detect(row.names(white_ci), pattern="0-15"),][[1]],2), 
                                         Upper_CI = round(white_ci[str_detect(row.names(white_ci), pattern="0-15"),][[2]],2)),
                                 c(Group = "40-54 (White only; adjusted by sex)", 
                                         OR = round(white_ecm[str_detect(names(white_ecm), pattern="40-54")][[1]],2), 
                                         Lower_CI = round(white_ci[str_detect(row.names(white_ci), pattern="40-54"),][[1]],2), 
                                         Upper_CI = round(white_ci[str_detect(row.names(white_ci), pattern="40-54"),][[2]],2)),
                                 c(Group = "55-69 (White only; adjusted by sex)", 
                                         OR = round(white_ecm[str_detect(names(white_ecm), pattern="55-69")][[1]],2), 
                                         Lower_CI = round(white_ci[str_detect(row.names(white_ci), pattern="55-69"),][[1]],2), 
                                         Upper_CI = round(white_ci[str_detect(row.names(white_ci), pattern="55-69"),][[2]],2)),
                                 c(Group = "70-84 (White only; adjusted by sex)", 
                                         OR = round(white_ecm[str_detect(names(white_ecm), pattern="70-84")][[1]],2), 
                                         Lower_CI = round(white_ci[str_detect(row.names(white_ci), pattern="70-84"),][[1]],2), 
                                         Upper_CI = round(white_ci[str_detect(row.names(white_ci), pattern="70-84"),][[2]],2)),
                                 c(Group = "85-90 (White only; adjusted by sex)", 
                                         OR = round(white_ecm[str_detect(names(white_ecm), pattern="85-90")][[1]],2), 
                                         Lower_CI = round(white_ci[str_detect(row.names(white_ci), pattern="85-90"),][[1]],2), 
                                         Upper_CI = round(white_ci[str_detect(row.names(white_ci), pattern="85-90"),][[2]],2))
                                 )
                           )

# checked sex coefficient was the same between the two interaction models
```

### Adjusted incidence rate ratios: modelling interaction between age and ethnicity using negative binomial

```{r interaction2}
### I checked for overdispersion in the final poisson model (ratio of residual deviance to degrees of freedom for the residuals should be 1), which may cause incorrect small standard errors and p-values of model coefficients. > 1 means overdispersion; Cameron and Trivedi 1990 suggest greater than 2 may require statistical intervention)

# poisson.model_ethnicity3$deviance/poisson.model_ethnicity3$df.residual gave slightly over 2, so I will compare to model fitted with negative binomial.
# plot(poisson.model_ethnicity3) residuals/predicted and root-standardised Pearson residuals plots weren't too bad, QQ plot looked fine.

nb.model_ethnicity5 <- glm.nb(formula = Cases_for_calculations ~ sex + 
                                     ethnicity_binary * age_group + offset(log(combined_denominator)),
                         data = adjusted_inc_rates_by_ethnicity, 
                         maxit = 1000)

# I initially got the following warning message so added maxit=1000:
# In glm.nb(formula = Cases_for_calculations ~ sex + ethnicity_binary *  :
#  alternation limit reached

summary(nb.model_ethnicity5)

# nb.model_ethnicity5$deviance/nb.model_ethnicity5$df.residual gave slightly less than 2
# plot(nb.model_ethnicity5) did not look very different to the poisson model.

ecm <- exp(coef(nb.model_ethnicity5))
ci <- exp(confint(nb.model_ethnicity5))

# Remove interaction coefficients to get only the adjusted age- and sex-specific coefficients for the White individuals
sensitivity_ecm <- ecm[!grepl(":", names(ecm))]
sensitivity_ci <- ci[!grepl(":", rownames(ci)),]

collated_ORs <- data.frame(rbind(collated_ORs,
                                       c(Group = "Female (NB adjusted by age * ethnicity)", 
                                         OR = round(sensitivity_ecm[str_detect(names(sensitivity_ecm), pattern="Female")][[1]],2), 
                                         Lower_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="Female"),][[1]],2), 
                                         Upper_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="Female"),][[2]],2)),
                                 c(Group = "0-15 (NB non-White only; adjusted by sex)", 
                                         OR = round(sensitivity_ecm[str_detect(names(sensitivity_ecm), pattern="0-15")][[1]],2), 
                                         Lower_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="0-15"),][[1]],2), 
                                         Upper_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="0-15"),][[2]],2)),
                                 c(Group = "40-54 (NB non-White only; adjusted by sex)", 
                                         OR = round(sensitivity_ecm[str_detect(names(sensitivity_ecm), pattern="40-54")][[1]],2), 
                                         Lower_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="40-54"),][[1]],2), 
                                         Upper_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="40-54"),][[2]],2)),
                                 c(Group = "55-69 (NB non-White only; adjusted by sex)", 
                                         OR = round(sensitivity_ecm[str_detect(names(sensitivity_ecm), pattern="55-69")][[1]],2), 
                                         Lower_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="55-69"),][[1]],2), 
                                         Upper_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="55-69"),][[2]],2)),
                                 c(Group = "70-84 (NB non-White only; adjusted by sex)", 
                                         OR = round(sensitivity_ecm[str_detect(names(sensitivity_ecm), pattern="70-84")][[1]],2), 
                                         Lower_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="70-84"),][[1]],2), 
                                         Upper_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="70-84"),][[2]],2)),
                                 c(Group = "85-90 (NB non-White only; adjusted by sex)", 
                                         OR = round(sensitivity_ecm[str_detect(names(sensitivity_ecm), pattern="85-90")][[1]],2), 
                                         Lower_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="85-90"),][[1]],2), 
                                         Upper_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="85-90"),][[2]],2))
                                 )
                           )

#then I repeated to collect the White coefficients
nb.model_ethnicity6 <- glm.nb(formula = Cases_for_calculations ~ sex + 
                                     ethnicity_binary2 * age_group + offset(log(combined_denominator)),
                         data = adjusted_inc_rates_by_ethnicity, 
                         maxit = 1000)

ecm <- exp(coef(nb.model_ethnicity6))
ci <- exp(confint(nb.model_ethnicity6))

sensitivity_ecm <- ecm[!grepl(":", names(ecm))]
sensitivity_ci <- ci[!grepl(":", rownames(ci)),]

collated_ORs <- data.frame(rbind(collated_ORs,
                                 c(Group = "0-15 (NB White only; adjusted by sex)", 
                                         OR = round(sensitivity_ecm[str_detect(names(sensitivity_ecm), pattern="0-15")][[1]],2), 
                                         Lower_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="0-15"),][[1]],2), 
                                         Upper_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="0-15"),][[2]],2)),
                                 c(Group = "40-54 (NB White only; adjusted by sex)", 
                                         OR = round(sensitivity_ecm[str_detect(names(sensitivity_ecm), pattern="40-54")][[1]],2), 
                                         Lower_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="40-54"),][[1]],2), 
                                         Upper_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="40-54"),][[2]],2)),
                                 c(Group = "55-69 (NB White only; adjusted by sex)", 
                                         OR = round(sensitivity_ecm[str_detect(names(sensitivity_ecm), pattern="55-69")][[1]],2), 
                                         Lower_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="55-69"),][[1]],2), 
                                         Upper_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="55-69"),][[2]],2)),
                                 c(Group = "70-84 (NB White only; adjusted by sex)", 
                                         OR = round(sensitivity_ecm[str_detect(names(sensitivity_ecm), pattern="70-84")][[1]],2), 
                                         Lower_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="70-84"),][[1]],2), 
                                         Upper_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="70-84"),][[2]],2)),
                                 c(Group = "85-90 (NB White only; adjusted by sex)", 
                                         OR = round(sensitivity_ecm[str_detect(names(sensitivity_ecm), pattern="85-90")][[1]],2), 
                                         Lower_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="85-90"),][[1]],2), 
                                         Upper_CI = round(sensitivity_ci[str_detect(row.names(sensitivity_ci), pattern="85-90"),][[2]],2))
                                 )
                           )
```

```{R export}
#Print table
collated_ORs %>%
  kable(booktabs=TRUE, align="l", col.names = c("Group", 
                       "Incidence rate ratio", 
                       "95% CI lower bound",
                       "95% CI upper bound"))

#write.csv(collated_ORs, "ccu075_01_M314_inc_ratios_ethnicity_age_sex_20250127.csv")

# Contextual information
# Adjusted incidence rate ratios for incidence of Takayasu arteritis by demographic group. 
# 
# Column 1: Demographic group relating to the specific incidence rate ratio, with details of the model adjustments in parentheses.
# Column 2: Incidence rate ratio of diagnosis with Takayasu arteritis within each demographic group (1 indicates reference group)
# Column 3: Poisson 95% confidence interval lower bound
# Column 4: Poisson 95% confidence interval upper bound
```

### Likelihood ratio test (added 18/02/2025)

```{r lrtest}
## The following is to extract the p value for interaction using lrtest
pval_model3v2 <- round(lrtest(poisson.model_ethnicity3, poisson.model_ethnicity2)$`Pr(>Chisq)`[[2]],2)

# Compare to a simpler model i.e., sex + ethnicity_binary + age_group
poisson.model_ethnicity5 <- glm(formula = Cases_for_calculations ~ sex + ethnicity_binary + age_group, 
                         offset = log(combined_denominator),
                         family = poisson(link="log"), 
                         data = adjusted_inc_rates_by_ethnicity)

summary(poisson.model_ethnicity5)

# Get coefficients
ecm <- exp(coef(poisson.model_ethnicity5)) # age- and sex-specific IRRs are remarkably consistent
ci <- exp(confint(poisson.model_ethnicity5))

options(scipen=999) # to turn off scientific notation; reset using scipen=0

pval_model3v5 <- round(lrtest(poisson.model_ethnicity3, poisson.model_ethnicity5)$`Pr(>Chisq)`[[2]],4)

pvals <- data.frame(rbind(c("Interaction model versus sex + ethnicity + age_group", pval_model3v2), 
                                 c("Interaction model versus sex + ethnicity_binary + age_group", pval_model3v5)
                                
))

pvals %>%
  kable(booktabs=TRUE, align="l", col.names = c("Comparison",
                                                "P value"))

# write.csv(pvals, "ccu075_01_lrtest_pvals_20250218.csv")

# Contextual information
# P values from likelihood ratio tests comparing different parameterisations of Poisson models for Takayasu arteritis incidence rate ratios. Interaction model = sex + ethnicity_binary + age_group
# 
# Column 1: Description of the comparison
# Column 2: P value
# 
# This export does not contain count or identifiable data

##Then extract the additional coefficients:
collated_ORs2 <- data.frame(rbind(c(Group = "Female (adjusted by age group and binary ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="Female")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="Female"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="Female"),][[2]],2)),
                                 c(Group = "0-15 (adjusted by sex and binary ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="0-15")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="0-15"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="0-15"),][[2]],2)),
                                 c(Group = "40-54 (adjusted by sex and binary ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="40-54")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="40-54"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="40-54"),][[2]],2)),
                                 c(Group = "55-69 (adjusted by sex and binary ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="55-69")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="55-69"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="55-69"),][[2]],2)),
                                 c(Group = "70-84 (adjusted by sex and binary ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="70-84")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="70-84"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="70-84"),][[2]],2)),
                                 c(Group = "85-90 (adjusted by sex and binary ethnicity)", 
                                         OR = round(ecm[str_detect(names(ecm), pattern="85-90")][[1]],2), 
                                         Lower_CI = round(ci[str_detect(row.names(ci), pattern="85-90"),][[1]],2), 
                                         Upper_CI = round(ci[str_detect(row.names(ci), pattern="85-90"),][[2]],2))
))

collated_ORs2 %>%
  kable(booktabs=TRUE, align="l", col.names = c("Group", 
                       "Incidence rate ratio", 
                       "95% CI lower bound",
                       "95% CI upper bound"))

#write.csv(collated_ORs2, "ccu075_01_M314_inc_ratios_binary_ethnicity_age_sex_20250218.csv")

# Contextual information
# Adjusted incidence rate ratios for incidence of Takayasu arteritis by demographic group. 
# 
# Column 1: Demographic group relating to the specific incidence rate ratio, with details of the model adjustments in parentheses.
# Column 2: Incidence rate ratio of diagnosis with Takayasu arteritis within each demographic group
# Column 3: Poisson 95% confidence interval lower bound
# Column 4: Poisson 95% confidence interval upper bound
```

### Version

```{r version, echo=FALSE}
sessionInfo()
```